\section{Tighter Strang Bound via Norm-of-Difference}
\label{apd:tighter}

We show that the Strang commutator-scaling bound can be tightened by
delaying the triangle inequality.
The standard bound
(\cref{sec:strang}, following Childs et al.~\cite{childsTheoryTrotterError2021})
uses a \emph{sum of norms}:
\begin{equation}\label{eq:strang-sum}
  \norm{S_2(t) - e^{tH}}
    \le \frac{\norm{\comm{B}{\comm{B}{A}}}}{12}\,t^3
      + \frac{\norm{\comm{A}{\comm{A}{B}}}}{24}\,t^3.
\end{equation}
We prove a bound using the \emph{norm of a difference}, which is always
$\le$ the sum and strictly tighter when the two double commutators
partially cancel.

\subsection{Setup}

Write $A' = A/2$, $H = A + B$, and recall the Strang splitting
$S_2(t) = e^{tA'}\,e^{tB}\,e^{tA'}$.
The Duhamel approach (\cref{sec:commutator}) gives
\begin{equation}\label{eq:strang-duhamel}
  S_2(t) - e^{tH} = e^{tH} \int_0^t \mathcal{T}_2(\tau)\,d\tau
  \qquad \text{(anti-Hermitian case: $\norm{e^{sX}} = 1$)},
\end{equation}
where $\mathcal{T}_2(\tau) = \text{Part}_A(\tau) + \text{Part}_B(\tau)$
is the Strang residual from the 4-factor product rule
(\cref{sec:strang}).

\subsection{Extracting the leading order}

The two parts have the following structure
(from the double FTC, \cref{sec:commutator}):

\paragraph{Part~A (from $A'$-conjugation).}
Define $g_1(s) = e^{sA'}\,\comm{A'}{B}\,e^{-sA'}$.
Then $\text{Part}_A(\tau) = \int_0^\tau (g_1(s) - g_1(\tau))\,ds$
(the ``subtract-constant-at-$\tau$'' trick).

Expanding $g_1(s) - g_1(\tau) = -\int_s^\tau g_1'(u)\,du$
where $g_1'(u) = e^{uA'}\,\comm{A'}{\comm{A'}{B}}\,e^{-uA'}$:
\begin{align}
  g_1(s) - g_1(\tau)
    &= -(\tau - s)\,\comm{A'}{\comm{A'}{B}} - \int_s^\tau
        \bigl(g_1'(u) - \comm{A'}{\comm{A'}{B}}\bigr)\,du
        \notag \\
    &= -(\tau - s)\,\comm{A'}{\comm{A'}{B}} + O(\tau^2).
        \label{eq:g1-expand}
\end{align}
The second term is bounded via the single FTC applied to $g_1'$:
\[
  g_1'(u) - \comm{A'}{\comm{A'}{B}}
    = \int_0^u e^{vA'}\,\comm{A'}{\comm{A'}{\comm{A'}{B}}}\,e^{-vA'}\,dv,
\]
so $\norm{g_1'(u) - \comm{A'}{\comm{A'}{B}}} \le
  \norm{\comm{A'}{\comm{A'}{\comm{A'}{B}}}}\,|u|$
(anti-Hermitian isometry).

Integrating \eqref{eq:g1-expand}:
\begin{equation}\label{eq:partA-expand}
  \text{Part}_A(\tau) = -\frac{\tau^2}{2}\,\comm{A'}{\comm{A'}{B}}
    + R_A(\tau),
\end{equation}
where $R_A(\tau)$ is a triple integral satisfying
\begin{equation}\label{eq:RA-bound}
  \norm{R_A(\tau)} \le
    \frac{\norm{\comm{A'}{\comm{A'}{\comm{A'}{B}}}}}{3}\,\tau^3.
\end{equation}
The bound \eqref{eq:RA-bound} follows from
$\norm{R_A(\tau)} \le \int_0^\tau \int_s^\tau |u|\,du\,ds
  \cdot \norm{\comm{A'}{\comm{A'}{\comm{A'}{B}}}}
  = \frac{\tau^3}{3}\,\norm{\comm{A'}{\comm{A'}{\comm{A'}{B}}}}$.

\paragraph{Part~B (from $B$-conjugation).}
The double-FTC remainder
$R_2(\tau) = e^{\tau B}\,A'\,e^{-\tau B} - A' - \tau\,\comm{B}{A'}$
satisfies (by the double-integral representation):
\begin{equation}\label{eq:R2-expand}
  R_2(\tau) = \frac{\tau^2}{2}\,\comm{B}{\comm{B}{A'}} + R_B(\tau),
  \qquad
  \norm{R_B(\tau)} \le
    \frac{\norm{\comm{B}{\comm{B}{\comm{B}{A'}}}}}{6}\,\tau^3.
\end{equation}
The bound on $R_B$ is exactly our triple-FTC result
(\texttt{norm\_exp\_conj\_sub\_comm2\_le\_of\_skewAdjoint}
in \leansrc{HigherCommutator.lean}).

Since $\text{Part}_B(\tau) = e^{\tau A'}\,R_2(\tau)\,e^{-\tau A'}$,
conjugation by $e^{\tau A'}$ introduces an additional correction:
\begin{equation}\label{eq:partB-expand}
  \text{Part}_B(\tau) = \frac{\tau^2}{2}\,\comm{B}{\comm{B}{A'}}
    + \underbrace{\frac{\tau^2}{2}\bigl(
      e^{\tau A'}\comm{B}{\comm{B}{A'}}e^{-\tau A'}
      - \comm{B}{\comm{B}{A'}}\bigr)}_{R_C(\tau)}
    + e^{\tau A'}\,R_B(\tau)\,e^{-\tau A'},
\end{equation}
where $R_C(\tau)$ is bounded by the single FTC on $\comm{B}{\comm{B}{A'}}$:
\begin{equation}\label{eq:RC-bound}
  \norm{R_C(\tau)} \le
    \frac{\norm{\comm{A'}{\comm{B}{\comm{B}{A'}}}}}{2}\,\tau^3.
\end{equation}

\subsection{The tighter bound}

Combining \eqref{eq:partA-expand} and \eqref{eq:partB-expand}:
\begin{equation}\label{eq:T2-refined}
  \mathcal{T}_2(\tau) = \frac{\tau^2}{2}\,D + R(\tau),
\end{equation}
where the \emph{effective double commutator} is
\begin{equation}\label{eq:D-def}
  \boxed{D \;=\; \comm{B}{\comm{B}{A'}} - \comm{A'}{\comm{A'}{B}}}
\end{equation}
and $R(\tau) = R_A(\tau) + R_C(\tau) + e^{\tau A'}\,R_B(\tau)\,e^{-\tau A'}$
satisfies
\begin{equation}\label{eq:R-bound}
  \norm{R(\tau)} \le T \cdot \tau^3,
  \qquad
  T = \frac{\norm{\comm{A'}{\comm{A'}{\comm{A'}{B}}}}}{3}
    + \frac{\norm{\comm{A'}{\comm{B}{\comm{B}{A'}}}}}{2}
    + \frac{\norm{\comm{B}{\comm{B}{\comm{B}{A'}}}}}{6}.
\end{equation}

Integrating:
\begin{align}
  \norm{S_2(t) - e^{tH}}
    &\le \int_0^t \norm{\mathcal{T}_2(\tau)}\,d\tau
    \le \int_0^t \Bigl(\frac{\norm{D}}{2}\,\tau^2 + T\,\tau^3\Bigr)\,d\tau
    \notag \\
    &= \frac{\norm{D}}{6}\,t^3 + \frac{T}{4}\,t^4.
    \label{eq:strang-tighter}
\end{align}

Substituting $A' = A/2$ (using $\comm{cX}{Y} = c\,\comm{X}{Y}$ for
central scalars, so $\comm{A'}{\comm{A'}{B}} = \tfrac{1}{4}\comm{A}{\comm{A}{B}}$
and $\comm{B}{\comm{B}{A'}} = \tfrac{1}{2}\comm{B}{\comm{B}{A}}$):

\begin{equation}\label{eq:strang-tighter-final}
  \boxed{
  \norm{S_2(t) - e^{tH}}
    \le \frac{1}{6}\left\|\frac{\comm{B}{\comm{B}{A}}}{2}
      - \frac{\comm{A}{\comm{A}{B}}}{4}\right\| t^3
    + T'\,t^4
  }
\end{equation}
where $T'$ involves triple commutators of $A$ and $B$ (see \eqref{eq:R-bound}
with $A' = A/2$).

\subsection{Comparison with the standard bound}

\paragraph{Relation.}
The standard bound \eqref{eq:strang-sum} can be written as
\[
  \frac{\norm{\comm{B}{\comm{B}{A}}}}{12}
  + \frac{\norm{\comm{A}{\comm{A}{B}}}}{24}
  = \frac{1}{6}\left(
    \frac{\norm{\comm{B}{\comm{B}{A}}}}{2}
    + \frac{\norm{\comm{A}{\comm{A}{B}}}}{4}
    \right).
\]
By the triangle inequality:
\[
  \left\|\frac{\comm{B}{\comm{B}{A}}}{2}
    - \frac{\comm{A}{\comm{A}{B}}}{4}\right\|
  \le
  \frac{\norm{\comm{B}{\comm{B}{A}}}}{2}
    + \frac{\norm{\comm{A}{\comm{A}{B}}}}{4},
\]
so the leading term of \eqref{eq:strang-tighter-final} is always
$\le$ the standard bound \eqref{eq:strang-sum}.
The improvement is strict whenever the two double commutators are
``aligned'' (i.e., partially cancel in the difference).

\paragraph{When is the tightening significant?}
The $O(t^4)$ correction in \eqref{eq:strang-tighter-final} is controlled
by triple commutator norms. For $t$ below the threshold
\[
  t^* = \frac{2}{3}\,
  \frac{\norm{\comm{B}{\comm{B}{A}}}/2 + \norm{\comm{A}{\comm{A}{B}}}/4
    - \norm{D}}{T'},
\]
the new bound \eqref{eq:strang-tighter-final} is tighter than
\eqref{eq:strang-sum}. In the Trotter regime ($t = O(1/n)$ with
$n \to \infty$), the $O(t^4)$ correction is negligible and the tightening
from the leading term dominates.

\paragraph{Example: symmetric Hamiltonians.}
When $A$ and $B$ enter symmetrically (e.g., even/odd sublattice
decomposition of a nearest-neighbor Hamiltonian), we may have
$\comm{B}{\comm{B}{A}} \approx -2\,\comm{A}{\comm{A}{B}}$
(up to boundary terms). In this case:
\[
  \frac{\comm{B}{\comm{B}{A}}}{2} - \frac{\comm{A}{\comm{A}{B}}}{4}
  \approx \frac{\comm{B}{\comm{B}{A}}}{2}
    + \frac{\comm{B}{\comm{B}{A}}}{8}
  = \frac{5}{8}\,\comm{B}{\comm{B}{A}},
\]
while the standard bound gives
$\norm{\comm{B}{\comm{B}{A}}}/2 + \norm{\comm{A}{\comm{A}{B}}}/4
  \approx \norm{\comm{B}{\comm{B}{A}}}$.
The tightening ratio is $5/8$, i.e., a $37.5\%$ reduction in the
leading coefficient.

\subsection{Why this tightening is absent from Childs et al.}
\label{sec:why-absent}

The standard bound \eqref{eq:strang-sum} appears in
Childs et al.~\cite{childsTheoryTrotterError2021} (Proposition~6) and
throughout the product formula literature. Several reasons explain why
the norm-of-difference formulation \eqref{eq:strang-tighter-final} has
not appeared previously.

\paragraph{Proof architecture.}
Childs et al.\ use a ``conjugation expansion'' (their Theorem~5) that
expands the error at each exponential interface independently, producing
separate terms for the $A$-$B$ and $B$-$A$ boundaries. The triangle
inequality on the resulting sum is the most direct bound within this
framework. The norm-of-difference requires an additional step:
recombining the interface terms, factoring out the common $\tau^2/2$
prefactor, and bounding the remainder with a \emph{triple}-commutator
estimate (\cref{eq:R-bound}). This extra step does not arise naturally
from the interface-by-interface analysis.

\paragraph{Framework compatibility.}
Childs et al.\ build a systematic framework around the
\emph{nested-commutator sum}
$\tilde\alpha_{\mathrm{comm}} =
  \sum \norm{\comm{H_{\gamma_{p+1}}}{\ldots\comm{H_{\gamma_2}}{H_{\gamma_1}}}}$,
which extends uniformly to any order~$p$ and any number of operators.
The effective commutator $D = \comm{B}{\comm{B}{A'}} - \comm{A'}{\comm{A'}{B}}$
is a \emph{difference} of two nested commutators---it does not fit into
the $\tilde\alpha_{\mathrm{comm}}$ framework as a single nested commutator.

\paragraph{Simplicity.}
The standard bound \eqref{eq:strang-sum} is a clean $t^3$ expression
with no correction term, yielding a simple Trotter-number formula
$r \sim (C\,t^3/\varepsilon)^{1/2}$ for resource estimation.
The tighter bound \eqref{eq:strang-tighter-final} introduces an $O(t^4)$
correction involving triple commutators, making the resource formula
more complex. For asymptotic scaling analysis, the simpler expression
is often preferred.

\paragraph{When the tightening matters.}
Despite these considerations, the norm-of-difference bound is
\emph{strictly more informative} for concrete resource estimation.
Given a specific Hamiltonian $H = A + B$, the effective commutator
$D$ is computable at negligible cost compared to the quantum simulation
itself. In the Trotter regime ($t = O(1/n)$), the $O(t^4)$ correction is
subdominant and the leading-term tightening dominates.
The formalization-driven insight here is that the Duhamel approach---by
producing the residual $\mathcal{T}_2(\tau)$ as a \emph{single}
algebraic expression---makes the norm-of-difference formulation natural,
whereas the interface-by-interface expansion does not.

\subsection{Extension to higher-order formulas}

The same principle extends to the fourth-order Suzuki formula $S_4$.
Childs et al.~\cite{childsTheoryTrotterError2021} (Proposition~7) bound
the $S_4$ error as a sum of 8 separate 4-fold commutator norms:
\[
  \norm{S_4(t) - e^{tH}}
    \le \sum_{k=1}^{8} \alpha_k\,\norm{C_k}\,t^5,
\]
with numerical coefficients $\alpha_k \in [0.0047, 0.0284]$.
The norm-of-difference formulation replaces this with
\[
  \norm{S_4(t) - e^{tH}} \le \norm{E_5}\,t^5 + O(t^6),
  \qquad
  E_5 = \sum_{k=1}^{8} \alpha_k\,C_k,
\]
where $E_5$ is the single algebraic error expression.
Since $\norm{E_5} \le \sum_k |\alpha_k|\,\norm{C_k}$ by the triangle
inequality, this bound is always $\le$ the standard one and strictly
tighter whenever the 4-fold nested commutators $C_k$ partially cancel.
For lattice Hamiltonians with spatial locality, many such cancellations
occur due to the geometric structure of the interaction graph, making
the tightening practically significant.

\subsection{Formalization status}

The triple-FTC building block (\cref{eq:R2-expand},
\leansrc{HigherCommutator.lean}) is fully formalized with 0
\texttt{sorry} axioms. The tighter Strang bound
\eqref{eq:strang-tighter-final} and its $S_4$ extension are work in
progress.
