\section{Conclusion and Future Work}
\label{sec:conclusion}

We have presented the first machine-checked formalization of the Lie--Trotter
product formula, together with its higher-order extensions and
commutator-scaling error bounds, in Lean~4 with the \Mathlib library.
The development comprises over 11\,400 lines with zero \texttt{sorry} axioms
(and three BCH-interface axioms confined to the Suzuki~$S_4$ track) and covers:
\begin{itemize}
  \item First-order Lie--Trotter convergence at rate $O(1/n)$;
  \item Symmetric Strang splitting at rate $O(1/n^2)$;
  \item Multi-operator generalizations of both formulas;
  \item First- and second-order commutator-scaling bounds via a novel
    Duhamel/FTC technique;
  \item A tighter second-order Strang bound via norm-of-difference that
    captures signed cancellations invisible to the sum-of-norms form;
  \item A fourth-order Suzuki $S_4$ track with four CAS-assisted
    BCH-based bounds, including a strictly tighter replacement
    (9$\times$--64$\times$) of Childs's 2021 heuristic prefactors and an
    existential-$\delta$ finite-$t$ bound via an explicit $\tau^7$ correction.
\end{itemize}

The Duhamel approach to commutator scaling---avoiding ODE uniqueness in favor
of the fundamental theorem of calculus---may be of independent interest as a
proof strategy for related product formula bounds.

\paragraph{Future directions.}
Several extensions are natural:
\begin{enumerate}
  \item \emph{Mathlib contributions.}
    The exponential norm bounds (\leanref{ExpBounds.lean}{72}{norm\_exp\_le},
    \leanref{ExpBounds.lean}{88}{norm\_exp\_sub\_one\_le},
    \leanref{ExpBounds.lean}{129}{exp\_sub\_one\_sub\_bound\_real})
    are general-purpose lemmas suitable for upstreaming to \Mathlib.

  \item \emph{Closing the two transitive Lean-BCH-private axioms.}
    Lean-Trotter no longer carries any own theorem-level BCH-interface
    axioms. The Suzuki~$S_4$ track now depends transitively on two
    private axioms inside the companion
    \href{https://github.com/Jue-Xu/Lean-BCH}{Lean-BCH} project:
    the three-factor symmetric BCH quintic Taylor bridge
    (\texttt{symmetric\_bch\_quintic\_sub\_poly\_axiom}, supports L2/L3)
    and the new five-factor palindromic septic identification
    (\texttt{suzuki5\_log\_product\_septic\_at\_suzukiP\_axiom},
    supports L4). Both have full multi-week discharge roadmaps in the
    project's \texttt{claude/} directory. The order-4 Taylor-coefficient
    identity \texttt{bch\_iteratedDeriv\_s4Func\_order4} was retired
    earlier via a three-slice chain that composes Lean-BCH's two-factor
    quintic payoff with a generic Taylor-match-from-norm lemma and
    \Mathlib's \texttt{iteratedDeriv} API; a Trotter-native
    operator-algebra alternative (via
    \texttt{scripts/compute\_sumQuadCorr\_factored.py}) remains as a
    nice-to-have that would remove the transitive dependence on Lean-BCH
    for the order-4 identity specifically.

  \item \emph{Baker--Campbell--Hausdorff formula.}
    The three-factor symmetric BCH identities and the two-factor
    palindromic quintic remainder are already formalized and imported
    from the Lean-BCH companion project (fully sorry-free as of
    revision~\texttt{c71d8f2}). Extending to five-factor palindromic
    BCH at orders $t^5$ and $t^7$, or to the general $n$-factor case,
    would complete the BCH-interface chain and automatically eliminate
    the remaining Suzuki~$S_4$ axioms.

  \item \emph{Specialization to quantum systems.}
    Connecting our abstract bounds to concrete gate-count estimates for
    Hamiltonian simulation on quantum computers would bridge the gap between
    formal mathematics and verified quantum algorithms.

  \item \emph{Unbounded operators.}
    The original Trotter--Kato theorem applies to strongly continuous
    semigroups on Banach spaces, covering unbounded generators.
    Formalizing this extension would require substantial additional
    functional analysis infrastructure in \Mathlib.
\end{enumerate}

More broadly, our project demonstrates that substantial results in operator
analysis---involving non-commutative algebra, power series manipulation,
Fr\'{e}chet calculus, and Bochner integration---are within reach of current
proof assistant technology.
We hope this work encourages further formalization of the mathematical
foundations underlying quantum algorithms.
