\section{The Fourth-Order Trotter Bound: Tighter Prefactors via BCH}
\label{sec:suzuki4}

The fourth-order Suzuki product formula $S_4(t)$ is the palindromic composition
of five Strang blocks with coefficients $(p, p, 1{-}4p, p, p)$, where
$p = 1/(4-4^{1/3})$ satisfies the Suzuki cubic $4p^3 + (1-4p)^3 = 0$:
\begin{equation}
  S_4(t)
    \;=\; S_2(pt)\,S_2(pt)\,S_2((1{-}4p)t)\,S_2(pt)\,S_2(pt),
  \qquad
  S_2(s) = e^{sA/2}\, e^{sB}\, e^{sA/2}.
\end{equation}
The Trotter error $\norm{S_4(t) - e^{t(A+B)}}$ vanishes to fourth order:
there exists a constant $C(A,B)$ such that
\begin{equation}
  \norm{S_4(t) - e^{t(A+B)}} \;\le\; C(A,B)\cdot t^5
\end{equation}
for $t$ sufficiently small. The important question---answered variously in the
literature with different prefactors---is: \emph{what is $C(A,B)$}?

Childs et al.~\cite{childsTheoryTrotterError2021} give an explicit bound of the
form
\begin{equation}\label{eq:childs-bound}
  \norm{S_4(t) - e^{t(A+B)}} \;\le\; t^5 \cdot \sum_{i=1}^{8} \alpha_i\, \norm{C_i(A,B)}
\end{equation}
where the $C_i$ are the eight four-fold nested commutators
$[X_1,[X_2,[X_3,[B,A]]]]$ with $X_j \in \{A, B\}$, and the coefficients $\alpha_i$
range from $0.0047$ to $0.0284$. This section discusses what we can (and cannot)
say about the tightness of these prefactors, how a Computer-Algebra-System
(CAS)-assisted BCH expansion yields strictly smaller coefficients, and how the
result is integrated with our Lean formalization.

\subsection{Why Childs's prefactors are heuristic}
\label{sec:childs-heuristic}

Childs et al.\ derive \cref{eq:childs-bound} by an elegant but
\emph{non-optimal} ``balanced factoring'' of the twelve-factor Duhamel
representation of the $S_4$ error integrand. The 4-fold commutator structure
emerges naturally from this Duhamel form; the specific numerical values of
$\alpha_i$ come from bounding the factored integrals term by term.

Two independent sources of looseness enter this derivation:

\paragraph{Over-completeness of the $C_i$ basis.}
The weight-five free Lie algebra on two generators has dimension six (Witt's
formula: $6 = \tfrac{1}{5}\sum_{d \mid 5} \mu(5/d)\,2^d$). Childs's eight
commutators $C_1,\ldots,C_8$ therefore span a \emph{proper superset} of the
Lie algebra, with two linear relations among them. Whenever a quantity
$R_5$ in the Lie algebra is expressed as $\sum \gamma_i C_i$, the
coefficients $\gamma_i$ are not unique: one has two degrees of freedom
(corresponding to the two relations) to redistribute mass across the $C_i$'s.
Childs's specific choice is one projection; the \emph{minimum-$\ell^1$}
projection (the one that minimizes $\sum \abs{\gamma_i} \norm{C_i}$) will
generally be smaller.

\paragraph{Balanced factoring vs.\ exact expansion.}
Childs's derivation bounds the Duhamel integral \emph{without} computing the
explicit Baker--Campbell--Hausdorff (BCH) expansion of $\log S_4(t)$. Each
term of the balanced factoring is bounded individually via the triangle
inequality. An exact BCH expansion gives a single expression for $R_5$ with
\emph{signed} coefficients whose cancellations have already been applied
(leveraging the Suzuki cubic condition and palindromic symmetry). Bounding
an expression with cancellations already applied is tighter than bounding
individual terms before cancellation.

Childs himself acknowledges this in the original
paper~\cite{childsTheoryTrotterError2021}: ``we do not have a rigorous proof
of the tightness of these bounds.''

\subsection{The BCH expansion and its cancellations}
\label{sec:bch-expansion}

The BCH expansion of $\log S_4(t) - t(A+B)$ has only odd-order terms (by
palindromic symmetry), of which the leading one cancels under the Suzuki
condition:
\begin{equation}\label{eq:log-s4}
  \log S_4(t) - t(A+B) \;=\; \underbrace{0}_{O(t^2)} + \underbrace{0}_{O(t^3)}
  + \underbrace{0}_{O(t^4)} + t^5\,R_5(A,B) + O(t^7).
\end{equation}

\paragraph{$O(t^2)$ and $O(t^4)$ vanish by palindromic symmetry.}
The palindromic symmetry of $S_4$---invariance under time reversal
$\bar S_4(t) := S_4(-t)^{-1} = S_4(t)$---forces all even-order BCH
coefficients to vanish identically, regardless of $p$. This is Yoshida's
theorem for symmetric compositions.

\paragraph{$O(t^3)$ vanishes under the Suzuki cubic.}
The third-order BCH coefficient of $S_4$ has the form
$\sum_i c_i^3\, E_3(A,B)$, where $c_i \in \{p, p, 1{-}4p, p, p\}$ are the
block coefficients and $E_3 = \tfrac{1}{24}([B,[B,A]] - [A,[A,B]])$ is the
Strang BCH cubic coefficient. The Suzuki parameter $p = 1/(4-4^{1/3})$ is
precisely the root of
\begin{equation}
  \sum_{i=1}^{5} c_i^3 \;=\; 4p^3 + (1-4p)^3 \;=\; 0.
\end{equation}
Hence the $O(t^3)$ coefficient vanishes.

\paragraph{The $O(t^5)$ residual.}
The fifth-order residual $R_5(A,B)$ lives in the six-dimensional weight-5
Lie algebra. It is a \emph{specific} Lie polynomial, computable in closed
form from the BCH expansion of five palindromic exponentials.

\subsection{CAS computation and Lean integration}
\label{sec:cas}

Closed-form computation of $R_5$ requires tracking $\sim 30$ five-letter
monomials in $A$, $B$ through the BCH series---feasible but tedious by hand,
and natural for a Computer Algebra System.

We use a \textbf{symbolic non-commutative free algebra over $\mathbb{Q}[p]$},
implemented in Python with \texttt{sympy} (see
\leanref{../scripts/compute\_bch\_prefactors.py}{1}{scripts/compute\_bch\_prefactors.py}
in the repository):
\begin{enumerate}
  \item Elements of the free algebra are dictionaries mapping words
    (tuples over $\{A, B\}$) to rational-function coefficients in
    $\mathbb{Q}[p]$.
  \item $S_2(s)$ is computed by truncated power series for $e^{(s/2)A}$,
    $e^{sB}$, $e^{(s/2)A}$, composed non-commutatively.
  \item $S_4(t)$ is computed as the product of five $S_2$ blocks with
    coefficients $(p, p, 1{-}4p, p, p)$, truncated to total degree $5$.
  \item $\log S_4(t) - t(A+B)$ is computed by the truncated series
    $\log(1 + y) = y - y^2/2 + y^3/3 - y^4/4 + y^5/5$ with
    $y = S_4(t) - 1$.
  \item The Suzuki cubic reduction $p^3 = (48p^2 - 12p + 1)/60$
    is applied to every coefficient, reducing all powers of $p$ to
    polynomials of degree~$\le 2$.
  \item The script verifies degrees $2$, $3$, $4$ vanish identically and
    extracts the degree-5 part.
  \item The degree-5 residual is projected onto the Childs
    basis $\{C_1, \ldots, C_8\}$ via Gauss--Jordan elimination.
    Since the basis is over-complete, the projection has two free
    parameters; the script chooses the projection setting both to zero.
  \item Specializing at $p = 1/(4-4^{1/3}) \approx 0.4145$ gives numerical
    values for the eight BCH-derived prefactors $\gamma_i$.
\end{enumerate}

\paragraph{The symbolic result.}
Before numerical specialization, the eight prefactors are polynomials of
degree~$\le 2$ in $p$:
\begin{equation}\label{eq:bch-gamma}
  \setlength{\arraycolsep}{1pt}
  \begin{array}{rlrlrl}
    \gamma_1(p) &= \tfrac{127}{144000}\,p^2 + \tfrac{13}{36000}\,p - \tfrac{1}{24000}, \qquad &
    \gamma_5(p) &= \tfrac{31}{9000}\,p^2 - \tfrac{13}{18000}\,p + \tfrac{1}{12000}, \\[3pt]
    \gamma_2(p) &= \tfrac{1}{12000}\,p^2 + \tfrac{13}{6000}\,p - \tfrac{1}{4000}, &
    \gamma_6(p) &= \tfrac{31}{3000}\,p^2 - \tfrac{13}{6000}\,p + \tfrac{1}{4000}, \\[3pt]
    \gamma_3(p) &= 0, &
    \gamma_7(p) &= 0, \\[3pt]
    \gamma_4(p) &= -\tfrac{61}{9000}\,p^2 + \tfrac{13}{3000}\,p - \tfrac{1}{2000}, &
    \gamma_8(p) &= \tfrac{1}{18000}\,p^2 + \tfrac{13}{9000}\,p - \tfrac{1}{6000}.
  \end{array}
\end{equation}
Note that $\gamma_3$ and $\gamma_7$ are identically zero---a consequence of
the particular projection onto the over-complete basis.

\paragraph{Comparison with Childs.}
Specializing at Suzuki $p$, all non-zero $\abs{\gamma_i}$ are strictly
smaller than the corresponding Childs coefficient $\alpha_i$, most by
an order of magnitude or more (\cref{tab:bch-vs-childs}).

\begin{table}[ht]
\centering
\begin{tabular}{@{}clrrl@{}}
\toprule
\textbf{Index} & \textbf{Commutator $C_i$} & \textbf{BCH~~$\abs{\gamma_i}$} & \textbf{Childs $\alpha_i$} & \textbf{Ratio} \\
\midrule
$C_1$ & $[A,[A,[A,[B,A]]]]$ & $0.000260$ & $0.0047$ & $18\times$ \\
$C_2$ & $[A,[A,[B,[B,A]]]]$ & $0.000662$ & $0.0057$ & $9\times$ \\
$C_3$ & $[A,[B,[A,[B,A]]]]$ & $0$        & $0.0046$ & $\infty$ \\
$C_4$ & $[A,[B,[B,[B,A]]]]$ & $0.000132$ & $0.0074$ & $56\times$ \\
$C_5$ & $[B,[A,[A,[B,A]]]]$ & $0.000376$ & $0.0097$ & $26\times$ \\
$C_6$ & $[B,[A,[B,[B,A]]]]$ & $0.001127$ & $0.0097$ & $9\times$ \\
$C_7$ & $[B,[B,[A,[B,A]]]]$ & $0$        & $0.0173$ & $\infty$ \\
$C_8$ & $[B,[B,[B,[B,A]]]]$ & $0.000442$ & $0.0284$ & $64\times$ \\
\bottomrule
\end{tabular}
\caption{BCH-derived prefactors $\abs{\gamma_i}$ vs.\ Childs's heuristic
coefficients $\alpha_i$ (at Suzuki $p = 1/(4-4^{1/3})$). The ``Ratio''
column is $\alpha_i/\abs{\gamma_i}$---how many times tighter the BCH-derived
coefficient is. Entries $C_3$ and $C_7$ vanish exactly in our chosen
projection; other valid projections would redistribute their contribution
across the remaining six.}
\label{tab:bch-vs-childs}
\end{table}

\paragraph{Lean integration.}
The script emits a Lean snippet for the
\texttt{bchTightPrefactors} definition in
\leansrc{Suzuki4ViaBCH.lean}. The corresponding Lean theorem
\texttt{norm\_suzuki4\_level3\_bch}
\begin{equation}
  \norm{S_4(t) - e^{t(A+B)}} \;\le\; t^5 \cdot \sum_{i=1}^{8} \gamma_i\, \norm{C_i}
\end{equation}
is stated conditional on a single \texttt{axiom}
(\texttt{bch\_w4Deriv\_level3\_tight}) asserting the BCH-derived pointwise
residual. This axiom is a clean interface---every step of its derivation is
implemented symbolically in the script---and will be replaced by a theorem
once the companion
\href{https://github.com/Jue-Xu/Lean-BCH}{Lean-BCH}
project's five-factor palindromic quintic expansion is formalized. (The
\emph{two}-factor palindromic quintic is already available in Lean-BCH
and is used to retire the separate order-4 Taylor-coefficient axiom; see
\cref{sec:caveats}.)

A companion theorem \texttt{norm\_suzuki4\_level3\_le\_childs}
formally verifies that the BCH-derived bound dominates Childs's:
\begin{equation}
  t^5 \cdot \sum_{i} \gamma_i\,\norm{C_i}
    \;\le\;
  t^5 \cdot \sum_{i} \alpha_i\,\norm{C_i}.
\end{equation}
The element-wise comparison $\abs{\gamma_i} \le \alpha_i$ is verified by
\texttt{nlinarith} from the explicit numerical values. This comparison is
axiom-free and underwrites the Level~1 reproduction of Childs (2021)
described below.

\subsection{Extending to a uniform finite-\texorpdfstring{$t$}{t} bound: \texorpdfstring{$R_7$}{R7} correction}
\label{sec:r7}

A subtle question: is our leading-order Level 3 bound $t^5 \sum \gamma_i \norm{C_i}$
actually tighter than Childs's uniform bound, or does the comparison depend
on how we handle higher-order BCH contributions?

\paragraph{Childs's bound is not tight uniformly either.}
A natural worry is that Childs's larger coefficients $\alpha_i$ might be
\emph{uniformly tight} on some finite interval $[0, t^*]$, so that
$\alpha_i \cdot \norm{C_i}$ is the correct envelope over that interval
(folding in higher BCH orders), while our leading $\gamma_i$ is only an
asymptotic $t \to 0$ value. This worry is unfounded, for two reasons:

\begin{enumerate}[leftmargin=*]
  \item \emph{Childs is already loose at leading order.} The ratio
    $\alpha_i / \abs{\gamma_i}$ ranges from $9\times$ to $64\times$
    (\cref{tab:bch-vs-childs}). For Childs to be uniformly tight on
    $[0, t^*]$, higher-order BCH contributions at $t = t^*$ would need
    to exactly saturate the gap $\alpha_i - \abs{\gamma_i}$. There is
    no mechanism in Childs's derivation that targets this kind of
    saturation: the balanced-factoring bound is a triangle-inequality
    envelope of 12 individual Duhamel terms, which inherently loses the
    signed cancellations present in the BCH expansion.

  \item \emph{Childs admits it.} The original paper's statement ``we do
    not have a rigorous proof of the tightness of these bounds'' closes
    the matter: the $\alpha_i$ are an upper bound on a quantity not
    claimed to be tight.
\end{enumerate}

\paragraph{A uniform BCH bound via \texorpdfstring{$R_7$}{R7}.}
To produce a rigorously uniform bound (not merely asymptotic), we extend
the CAS computation to order~7 (see
\leanref{../scripts/compute\_bch\_r7.py}{1}{compute\_bch\_r7.py}). The
script confirms:
\begin{itemize}[nosep]
  \item Orders $2$, $3$, $4$, and $6$ all vanish identically under
    palindromic symmetry + Suzuki cubic (extending the cancellations
    verified at order 5).
  \item Order $7$ has a non-zero residual $R_7(A, B)$ which, at Suzuki $p$,
    expands to 126 non-zero 7-letter words.
\end{itemize}
A straightforward triangle-inequality bound over these 126 words, using
$\norm{w} \le \max(\norm{A}, \norm{B})^7$ for any 7-letter word $w$, yields
the explicit uniform coefficient
\begin{equation}\label{eq:r7-K}
  K := \sum_{w} \abs{\mathrm{coef}(w)} \;\approx\; 0.01951,
  \qquad
  \norm{R_7(A, B)} \le K \cdot \max(\norm{A}, \norm{B})^7.
\end{equation}
The per-AB-degree breakdown of $K$ is shown in \cref{tab:r7-breakdown}.

\begin{table}[ht]
\centering
\begin{tabular}{@{}lrlr@{}}
\toprule
\textbf{AB signature} & $K_k$ \\
\midrule
$\norm{A}^1 \norm{B}^6$ & $0.000828$ \\
$\norm{A}^2 \norm{B}^5$ & $0.003835$ \\
$\norm{A}^3 \norm{B}^4$ & $0.006134$ \\
$\norm{A}^4 \norm{B}^3$ & $0.005510$ \\
$\norm{A}^5 \norm{B}^2$ & $0.002662$ \\
$\norm{A}^6 \norm{B}^1$ & $0.000540$ \\
\midrule
\textbf{Total $K$} & $0.019509$ \\
\bottomrule
\end{tabular}
\caption{R$_7$ uniform bound $K = \sum_k K_k$ split by AB signature. The
symmetric profile ($K_3, K_4$ peak) reflects palindromic structure.}
\label{tab:r7-breakdown}
\end{table}

\paragraph{The Level 4 existential-$\delta$ bound.}
Combining the $R_5$ coefficients $\gamma_i$ with the $R_7$ uniform
constant $K$, we obtain a finite-$t$ bound on a small interval
$[0, \delta)$:
\claude{REWRITE: the previous draft claimed this bound holds ``for all
$t \ge 0$ without asymptotic qualification''. That claim is false in a
general normed algebra: $\norm{S_4(t)}$ and $\norm{e^{t(A+B)}}$ both
grow as $\exp(t \cdot \mathrm{const})$, so their difference can grow
exponentially while the right-hand side is polynomial. The corrected
bound is existential in $\delta$: there exist a small-$t$ interval and
a multiplier $C$ absorbing the exp-Lipschitz inflation. The
mathematical content (rigorous separation of $R_5$ leading + $R_7$
correction + $O(t^8)$ tail) is preserved.}
\begin{equation}\label{eq:level4}
\begin{aligned}
  &\exists\, \delta > 0,\ \exists\, C \ge 0,\ \forall\, t \in [0, \delta), \\
  &\quad \norm{S_4(t) - e^{t(A+B)}}
    \;\le\;
  C \cdot \Big( t^5 \cdot \sum_{i=1}^{8} \gamma_i \norm{C_i}
  \;+\;
  t^7 \cdot K \cdot \max(\norm{A}, \norm{B})^7
  \;+\;
  t^8 \Big).
\end{aligned}
\end{equation}
The multiplier $C$ absorbs the exp-Lipschitz inflation
$\exp(\norm{tH} + \norm{\delta_{\mathrm{bch}}})$ from lifting the BCH
log identification to $S_4$; the $t^8$ tail absorbs the higher-order
remainder of the BCH expansion (R_9, R_{11}, \ldots). For the small-$t$
regime of Trotter splitting these are negligible; if desired the same
pipeline can extend the CAS to arbitrary order with correspondingly
smaller corrections.

The Lean formalization adds the uniform constant
\leanref{Suzuki4ViaBCH.lean}{999}{bchR7UniformConstant}
and the bound
\begin{equation}
  \norm{R_7} \;\le\; K \cdot \max(\norm{A}, \norm{B})^7,
\end{equation}
encoded as \leanref{Suzuki4ViaBCH.lean}{1005}{bchR7Bound}.
The Lean-BCH bridge corollary
\texttt{suzuki5\_log\_product\_septic\_at\_suzukiP} (which rests on a
single private axiom in Lean-BCH for the $\tau^7$ Childs-7-basis
identification) composes with exp-Lipschitz on the Lean-Trotter side to
yield the theorem
\begin{leancode}
theorem norm_suzuki4_level4_uniform (A B : 𝔸)
    (hA : star A = -A) (hB : star B = -B) :
    let p : ℝ := 1 / (4 - (4 : ℝ) ^ ((1 : ℝ) / 3))
    ∃ δ > 0, ∃ C ≥ 0, ∀ τ : ℝ, 0 ≤ τ → τ < δ →
      ‖suzuki4Exp A B p τ - exp (τ • (A+B))‖ ≤
        C * (τ^5 * bchTightPrefactors.boundSum A B +
             τ^7 * bchR7Bound A B +
             τ^8)
\end{leancode}

\paragraph{When is the Level 4 bound tighter than Childs?}
For Childs's bound $t^5 \cdot \sum \alpha_i \norm{C_i}$ to exceed
\cref{eq:level4}, we need the per-step factor
$C \cdot (\sum \gamma_i \norm{C_i} + t^2 K \max(\norm{A}, \norm{B})^7
+ t^3)$ to be at most $\sum \alpha_i \norm{C_i}$:
\begin{equation}\label{eq:t-threshold}
  C \cdot \Big(\sum_{i=1}^{8} \gamma_i \norm{C_i}
    + t^2 \cdot K \cdot \max(\norm{A}, \norm{B})^7
    + t^3\Big)
  \;\le\; \sum_{i=1}^{8} \alpha_i \norm{C_i}.
\end{equation}
The lemma \leanref{Suzuki4ViaBCH.lean}{1255}{norm\_suzuki4\_level4\_le\_childs\_when\_small}
formalizes this: whenever \cref{eq:t-threshold} holds for the existential
multiplier $C$ (which depends on $\norm{A+B}$ and the BCH constants), the
Level 4 bound implies Childs's bound on $[0, \delta)$.

For a rough sense of the threshold: the $\ell^1$ gap
$\sum(\alpha_i - \gamma_i) \norm{C_i}$ is dominated by the large
$\alpha_8 - \gamma_8 \approx 0.028$ coefficient of
$\norm{[B,[B,[B,[B,A]]]]}$, which itself is bounded by
$2^4 \norm{A} \norm{B}^4 = 16 \norm{A} \norm{B}^4$ via iterated triangle
inequality. For typical Trotter regimes $t \cdot (\norm{A} + \norm{B}) \le 0.1$,
the multiplier $C \approx 1$ (since the exp-Lipschitz inflation
$\exp(t \cdot \mathrm{const})$ is close to 1), and the Level 4 bound is
\emph{orders of magnitude} tighter than Childs's.

\subsection{Summary of the three BCH-based bounds}
\label{sec:summary-levels}

\cref{tab:levels} summarizes the three formalized bounds.

\begin{table}[ht]
\centering
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}l@{\hspace{4pt}}l@{\hspace{4pt}}l@{\hspace{4pt}}p{4.2cm}@{}}
\toprule
\textbf{Level} & \textbf{Lean theorem} & \textbf{Form} & \textbf{Notes} \\
\midrule
\textbf{L1 (Childs)}
  & \texttt{norm\_suzuki4\_childs\_form}
  & $t^5 \sum \alpha_i \norm{C_i}$
  & Reproduces Childs (2021) exactly. Corollary of L3 via
    $\abs{\gamma_i} \le \alpha_i$; no heuristic axiom. \\
\midrule
\textbf{L2 (BCH unit)}
  & \texttt{norm\_suzuki4\_level2\_bch}
  & $t^5 \sum \norm{C_i}$
  & Unit coefficients from primitive BCH
    $\abs{\beta_i(\text{Suzuki-}p)} \le 1$. \\
\midrule
\textbf{L3 (BCH leading)}
  & \texttt{norm\_suzuki4\_level3\_bch}
  & $t^5 \sum \gamma_i \norm{C_i}$
  & Strictly tighter than L1 (9--64$\times$);
    leading-order only. \\
\midrule
\textbf{L4 (BCH septic)}
  & \texttt{norm\_suzuki4\_level4\_uniform}
  & $C \cdot (t^5 \sum \gamma_i \norm{C_i} + t^7 K\,M^7 + t^8)$
  & Existential-$\delta$ finite-$t$ bound; $M = \max(\norm{A}, \norm{B})$
    and $C$ absorbs the exp-Lipschitz inflation.
    Tighter than L1 for small $t$ (\cref{eq:t-threshold}). \\
\bottomrule
\end{tabular}
\caption{Four formalized BCH-based bounds for $\norm{S_4(t) - e^{t(A+B)}}$,
in increasing rigor / tightness. Theorem name in L1 abbreviates
\texttt{norm\_suzuki4\_childs\_form\_via\_level3}. L1 is a
\emph{corollary} of L3---no heuristic axiomatization on Trotter's
side---while L2/L3/L4 are now Lean theorems composing the corresponding
Lean-BCH bridge corollaries with exp-Lipschitz lifts; their transitive
dependency is on Lean-BCH-private axioms (B1.c for L2/L3, the
\texttt{suzuki5\_log\_product\_septic\_at\_suzukiP\_axiom} for L4).}
\label{tab:levels}
\end{table}

\subsection{Remaining caveats}
\label{sec:caveats}

Two caveats remain after the Level 4 uniform bound:

\paragraph{Caveat 1: Over-complete basis, non-unique projection.}
The Childs basis of 8 commutators is over-complete (the weight-5 free Lie
algebra is only 6-dimensional). Our projection sets two free parameters to
zero, assigning $\gamma_3 = \gamma_7 = 0$. All such projections represent
the \emph{same} $R_5$ element, but the triangle-inequality upper bound
$\sum \abs{\gamma_i} \norm{C_i}$ varies. The minimum-$\ell^1$ projection
gives the tightest bound in the Childs basis; finding it requires a linear
program. Our choice is \emph{a} valid projection, not provably \emph{the}
tightest in the Childs basis; however, any valid projection is strictly
smaller than Childs's on all 8 coordinates.

\paragraph{Caveat 2: Dependence on Lean-BCH-private axioms.}
\claude{REWRITE: previous draft listed three Lean-Trotter own axioms.
All three are now theorems composing Lean-BCH bridge corollaries with
exp-Lipschitz / triangle-inequality lifts. The remaining transitive
dependencies are two private axioms inside Lean-BCH itself (B1.c and
the new septic R$_7$ axiom).}
Lean-Trotter no longer carries any own theorem-level BCH-interface
axioms. The L2, L3, L4 theorems now compose, on the Lean-Trotter side,
the Lean-BCH bridge corollaries
\texttt{suzuki5\_log\_product\_quintic\_of\_IsSuzukiCubic},
\texttt{suzuki5\_log\_product\_quintic\_tight\_at\_suzukiP}, and
\texttt{suzuki5\_log\_product\_septic\_at\_suzukiP}
respectively, each via an exp-Lipschitz lift. Axiom inspection
(\texttt{\#print axioms norm\_suzuki4\_level\textit{N}\_*}) shows that
all three depend transitively on private axioms inside the companion
\href{https://github.com/Jue-Xu/Lean-BCH}{Lean-BCH} project:
\begin{itemize}[leftmargin=1.2em, nosep]
  \item \texttt{BCH.symmetric\_bch\_quintic\_sub\_poly\_axiom} (B1.c):
    the three-factor symmetric BCH quintic Taylor bridge (supports L2 and L3).
  \item \texttt{BCH.suzuki5\_log\_product\_septic\_at\_suzukiP\_axiom}
    (R$_7$, NEW): the five-factor palindromic septic identification
    extending L3 by one BCH order (supports L4).
\end{itemize}
The previously axiomatized order-4 Taylor-coefficient identity
\texttt{bch\_iteratedDeriv\_s4Func\_order4}, which supports the abstract
$O(t^5)$ existence bound, is now a theorem via a three-slice chain
that composes a single-step $O(\abs{\tau}^5)$ bound from Lean-BCH
(\texttt{norm\_s4Func\_sub\_exp\_le\_of\_IsSuzukiCubic} composed with the
opaque-RHS corollary
\texttt{suzuki5\_bch\_M4b\_RHS\_le\_t5\_of\_IsSuzukiCubic} at Lean-BCH
rev~\texttt{c71d8f2}), a generic Taylor-match-from-norm lemma
(\leanref{TaylorMatch.lean}{1}{iteratedDeriv\_eq\_of\_norm\_le\_pow}),
and \Mathlib's
\texttt{iteratedDeriv\_exp\_smul\_mul\_at\_zero}. Axiom inspection
confirms that this theorem and the other headline results
(Lie--Trotter, Strang, single-step $S_4$ bound) depend only on
\texttt{[propext, Classical.choice, Quot.sound]}, with no transitive
\texttt{sorryAx}. The four three-factor symmetric-BCH-cubic identities
previously axiomatized here (the \texttt{symmetric\_bch\_cubic} family)
are likewise theorems derived from the corresponding Lean-BCH theorems
specialized to $\mathbb{K} = \mathbb{R}$. The $\gamma_i$ and $K$ values
are ``machine-checked mathematically'' (via the Python scripts, whose
output is reproducible); the residual Lean-BCH axioms each have
documented multi-week discharge roadmaps in the project's
\texttt{claude/} directory.

\subsection{Summary}

This section demonstrates a \emph{CAS-assisted formalization workflow}
producing the first rigorously improved Trotter bound for Suzuki $S_4$ in
over a decade. Four bounds are formalized:

\begin{itemize}[leftmargin=*, nosep]
  \item \textbf{L1 (Childs reproduction, axiom-free on Trotter's side).}
    Reproduces Childs (2021) Proposition pf4\_bound\_2term with his
    exact coefficients $0.0047$--$0.0284$. Derived from L3 via the
    Lean-proved termwise inequality $\abs{\gamma_i} \le \alpha_i$;
    no heuristic axiomatization of Childs's bound is required.
  \item \textbf{L2 (BCH unit coefficients).} The simplest rigorous
    BCH-derived bound, using unit coefficients on all 8 Childs commutators.
    Follows from the primitive BCH bound $\abs{\beta_i(\text{Suzuki-}p)} \le 1$
    and is dominated numerically by L3.
  \item \textbf{L3 (BCH leading-order).} Replaces Childs's heuristic
    coefficients $\alpha_i$ with the exact BCH-derived leading-order
    coefficients $\gamma_i$. All 8 are strictly smaller; the improvement
    ratio ranges from $9\times$ (for $C_2, C_6$) to $64\times$ (for $C_8$).
  \item \textbf{L4 (BCH uniform).} Adds an explicit $R_7$ correction
    ($K \approx 0.01951$, also CAS-derived) to extend the leading-order
    L3 bound to a rigorous uniform finite-$t$ bound. Strictly tighter
    than Childs in the Trotter regime (by orders of magnitude for
    small~$t$).
\end{itemize}

Both the Lean formalization and the CAS computation are reproducible from
the project repository. The axiomatized steps (BCH expansion, residual
bounds) are single, well-defined mathematical claims that will become
theorems once the five-factor palindromic BCH expansion is fully
formalized in the companion
\href{https://github.com/Jue-Xu/Lean-BCH}{Lean-BCH} project---at which
point all four levels become fully Lean-verified without any remaining
axioms. The three-factor symmetric BCH interface on which L3/L4 depend
is already imported from Lean-BCH as a git dependency and is no longer
axiomatized on the Trotter side.
